Resolvable Ambiguity

### 2019-07-08 paper/paper.tex
Polish text in 6.1

However, there is a case where this alternate representation is not fully isomorphic; namely when a pair of parentheses would yield a zero-length range. As an example, consider the two words ``\verb|(())|'' and ``\verb|()()|''. Both of these have an alternate representation of $(\text{``}\text{''}, \{\range{0}{0}, \range{0}{0}\})$. In versions~1 and 2 this corresponds to grouping parentheses surrounding zero-length productions. We can avoid this case by finding every production $N -> l : r$ where $\epsilon \in L(r)$, replacing it with $N -> l : r'$ where $L(r') = L(r) \setminus \{\epsilon\}$, and replacing every use of $N$ with $N + \epsilon$. The one exception is if the start symbol has a nullable production, but that can be trivially special cased. As such, going forward, we will assume no zero-length grouping parentheses.

@@

\begin{lemma}
  $\forall w'_1, w'_2 \in \words(t).\ \basic(w'_1) = \basic(w'_2)$, i.e., all $w'$ have the same basic word.
  \label{lemma:same-basic}
\end{lemma}
\noindent Intuitively, $\words$ first applies the ``inverse'' of $\semantic$, i.e., adding some number of grouping parentheses nodes between pre-existing nodes, then $\yield$, which flattens the tree to a word. The only terminals changed in this process are parentheses, i.e., different $w'$ can only differ in their range bags. This also implies that two trees that share a word (i.e., two trees $t_1$ and $t_2$ that have a word $w'$ such that $\parse(w') \supseteq \{t_1, t_2\}$) must also share a basic word.

\begin{lemma}
  $\exists R$ such that $\exists w' \in \words(t).\ \rangebag(w') = R$ and $\forall w' \in \words(t).\\ R \subseteq \support(\rangebag(w'))$, i.e., some parentheses are required.
  \label{lemma:required-parentheses}
\end{lemma}
\noindent For example, removing the parentheses in ``\verb|(1 + 2) * 3|'' changes the interpretations produced. Each required range is a direct consequence of a mark on a non-terminal in $D$. We will write $w'_\bot$ for the word implied by the first quantified expression. It is unique, since its basic word is fixed by $t$, and its rangebag is a set.

>>>>

However, there is a case where this alternate representation is not fully isomorphic; namely when a pair of parentheses would yield a zero-length range. As an example, consider the two words ``\verb|(())|'' and ``\verb|()()|''. Both of these have an alternate representation of $(\text{``}\text{''}, \{\range{0}{0}, \range{0}{0}\})$. In versions~1 and 2 this corresponds to grouping parentheses surrounding zero-length productions. We can avoid this case by finding every production $N -> l : r$ where $\epsilon \in L(r)$, replacing it with $N -> l : r'$ where $L(r') = L(r) \setminus \{\epsilon\}$, and replacing every use of $N$ with $N + \epsilon$. The one exception is if the start symbol has a nullable production, but that can be trivially special cased. Thus, we will assume no zero-length grouping parentheses in the following.

@@

\begin{lemma}
  $\forall w'_1, w'_2 \in \words(t).\ \basic(w'_1) = \basic(w'_2)$, i.e., all $w'$ have the same basic word.
  \label{lemma:same-basic}
\end{lemma}
\noindent Intuitively, $\words$ first applies the ``inverse'' of $\semantic$, i.e., adding some number of grouping parentheses nodes between pre-existing nodes, then $\yield$, which flattens the tree to a word. The only terminals changed in this process are parentheses, i.e., different $w'$ can only differ in their range bags. This also implies that two trees that share a word (i.e., trees $t_1$ and $t_2$ that have a word $w'$ such that $\parse(w') \supseteq \{t_1, t_2\}$) must also share a basic word.

\begin{lemma}
  $\exists R$ such that $\exists w' \in \words(t).\ \rangebag(w') = R$ and $\forall w' \in \words(t).\\ R \subseteq \support(\rangebag(w'))$, i.e., some parentheses are required.
  \label{lemma:required-parentheses}
\end{lemma}
\noindent For example, removing the parentheses in ``\verb|(1 + 2) * 3|'' changes the produced parse trees. Each required range is a direct consequence of a mark on a non-terminal in $D$. We will write $w'_\bot$ for the word implied by the first quantified expression. It is unique, since its basic word is fixed by $t$, and its rangebag is a set.
